---
title: "ML PROJECT"
output: html_document
---

There are several devices which can effectively collect the data in continuous storage everyday,many of them are related to fitbits and trackers which continuously monitor different parameters of health helping humans keep a check on their health.

In this project,we will use the data from accelerators on the belt of 6 participants. they were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

 The data consists of Training and Testing sets .
 
 The goal of the project is to predict the manner in which they did the exercise, this is the 'classe' variable in the training set.
 
 DATA LOADING AND PROCESSING
 ```{r}
 library(caret)
 library(rpart)
 library(rpart.plot)
 library(RColorBrewer)
library(rattle)
library(randomForest)
library(corrplot)
library(gbm)
```

GETTING,CLEANING AND EXPLORING THE DATA
```{r}
train_in<-read.csv('./pml-training.csv',header=T)
valid_in<-read.csv('./pml-testing.csv',header=T)
dim(valid_in)
```

As shown there are 19622 observations and 160 variables in the training dataset.

CLEANING THE INPUT DATA 

we remove the variables containing the missing values.then we display the dimensions of the displayed set.
```{r}
trainData<-train_in[,colSums(is.na(train_in))==0]
validData<-valid_in[,colSums(is.na(valid_in))==0]
dim(trainData)
dim(validData)
```

We now remove the forst 7 variables as they have little impact on the outcome class.
```{r}
trainData<-trainData[,-c(1:7)]
validData<-validData[,-c(1:7)]
dim(trainData)
dim(validData)
```

PREPARING THE DATASETS FOR PREDICTION

Preparing the data prediction by splitting the traingng data into 70% and 30% as test data. this splitting will serve also to compute the out of sample errors.

The test data renamed: valid_in (validate data) will stay as is and will be used later test the prediction algorithm on the 20 cases.
```{r}
set.seed(1234)
inTrain<-createDataPartition(trainData$classe,p=0.7,list=FALSE)
trainData<-trainData[inTrain,]
testData<-trainData[-inTrain,]
dim(trainData)
dim(testData)
```

CLEANING EVEN FURTHER BY REMOVING THE VARIABLES THAT ARE NEAR-ZERO-VARIANCE
```{r}
NZV<-nearZeroVar(trainData)
trainData<-trainData[,-NZV]
testData<-testData[,-NZV]
dim(trainData)
dim(testData)
```

After this cleaning we are down to 53 variables.
```{r}
cor_mat<-cor(trainData[,-53])
corrplot(cor_mat,order="FPC",method="color",type="upper",tl.cex=0.8,tl.col=rgb(0,0,0))
```

